### 2024-04-29 sections/5-discussion.tex
careful edit

We have presented CAKES, a set of three algorithms for fast $k$-NN search that are generic over a variety of distance functions.
CAKES's algorithms are exact when the distance function in use is a metric (see definition in~\ref{sec:methods}).
Even under Cosine distance, which is not a metric, CAKES's algorithms exhibit nearly perfect recall (i.e. always greater than $0.9995$).

@@

Figure~\ref{fig:results:scaling-plots} also shows that HNSW and ANNOY have near constant throughput as the multiplier increases, however, their recall continues to decrease as the multiplier increases.
This, coupled with the prohibitively increasing cost (in terms of memory and time) of building the indices for HNSW and ANNOY, makes these algorithms unsuitable for real-world applications in which the dataset cardinalities are growing exponentially.
CAKES's tree is orders of magnitude cheaper to build and CAKES's algorithms, while a little slower than HNSW and ANNOY, still show near constant scaling with cardinality in addition to having perfect recall.

When written with the same notation as used in Section~\ref{sec:methods:knn-search:complexity-of-sieve-methods}, we see that the time complexity of Repeated $\rho$-NN is $\mathcal{O}(\tfrac{\mathcal{T}}{d} + \mathcal{L})$, where $\mathcal{T}$ is the time complexity of tree-search and $\mathcal{L}$ is the time complexity of leaf-search. 
Even though Repeated $\rho$-NN has the lowest time complexity (compare to $\mathcal{O}(\mathcal{T}\log{(\mathcal{T}+\mathcal{L})} + \mathcal{L}\log{(\mathcal{T}+\mathcal{L})})$ for Breadth-First Sieve and $\mathcal{O}(\mathcal{T}\log{\mathcal{T}} + \mathcal{L}\log{k})$ for Depth-First Sieve) of the three CAKES algorithms, it is not always the fastest algorithm empirically. 
We believe that some of this discrepancy can be explained by the fact that if Repeated $\rho$-NN significantly ``overshoots'' the correct radius for $k$ hits ($\rho_k$) during tree-search, leaf-search will require looking at more points, rendering the true scaling factor higher than that in~\ref{eq:methods:repeated-rnn-complexity}.
This overshot can occur when the LFDs of clusters near the query are not concentrated around their expectation. 
For example, if most clusters near the query have very low LFD except for one anomaly with very high LFD, 
the harmonic mean LFD $\mu$ can still be low, so the factor of radial increase~\ref{eq:methods:repeated-rnn-factor} may be much larger than necessary for guaranteeing $k$ hits.  
This suggests that rather than using the reciprocal of the \textit{harmonic} mean LFD in~\ref{eq:methods:repeated-rnn-factor}, we may achieve better results with a mean that is more sensitive to high outliers, such as the geometric mean.
We leave it as an avenue for future work to characterize when Repeated $\rho$-NN significantly ``overshoots'' $\rho_k$ and to improve upon the factor of radial increase in~\ref{eq:methods:repeated-rnn-factor} so that this occurs less frequently and with less severity.

>>>>

We have presented CAKES, a set of three algorithms for fast $k$-NN search that are generic over a variety of distance functions.
CAKES's algorithms are exact when the distance function in use is a metric (see definition in~\ref{sec:methods}).
Even under cosine distance, which is not a metric, CAKES's algorithms exhibit nearly perfect recall (i.e. always greater than $0.9995$).

@@

Figure~\ref{fig:results:scaling-plots} also shows that HNSW and ANNOY have near constant throughput as the multiplier increases, however, their recall continues to decrease as the multiplier increases.
This, coupled with the prohibitively increasing cost (in terms of memory and time) of building the indices for HNSW and ANNOY, makes these algorithms unsuitable for real-world applications in which the dataset cardinalities are growing exponentially.
CAKES's tree is orders of magnitude faster and less memory-intensive to build and CAKES's algorithms, while a little slower than HNSW and ANNOY, still show near constant scaling with cardinality in addition to having perfect recall.

When written with the same notation as used in Section~\ref{sec:methods:knn-search:complexity-of-sieve-methods}, we see that the time complexity of Repeated $\rho$-NN is $\mathcal{O}(\tfrac{\mathcal{T}}{d} + \mathcal{L})$, where $\mathcal{T}$ is the time complexity of tree-search and $\mathcal{L}$ is the time complexity of leaf-search. 
Even though Repeated $\rho$-NN has the lowest time complexity (compare to $\mathcal{O}(\mathcal{T}\log{(\mathcal{T}+\mathcal{L})} + \mathcal{L}\log{(\mathcal{T}+\mathcal{L})})$ for Breadth-First Sieve and $\mathcal{O}(\mathcal{T}\log{\mathcal{T}} + \mathcal{L}\log{k})$ for Depth-First Sieve) of the three CAKES algorithms, it is not always the fastest algorithm empirically. 
We believe that some of this discrepancy can be explained by the fact that if Repeated $\rho$-NN significantly ``overshoots'' the correct radius for $k$ hits ($\rho_k$) during tree-search, leaf-search will require looking at more points, rendering the true scaling factor higher than that in Equation~\ref{eq:methods:repeated-rnn-complexity}.
This overshot can occur when the LFDs of clusters near the query are not concentrated around their expectation. 
For example, if most clusters near the query have very low LFD except for one anomaly with very high LFD, 
the harmonic mean LFD $\mu$ can still be low, so the factor of radial increase~\ref{eq:methods:repeated-rnn-factor} may be much larger than necessary for guaranteeing $k$ hits.  
This suggests that rather than using the reciprocal of the \textit{harmonic} mean LFD in~\ref{eq:methods:repeated-rnn-factor}, we may achieve better results with a mean that is more sensitive to high outliers, such as the geometric mean.
We leave it as an avenue for future work to characterize when Repeated $\rho$-NN significantly ``overshoots'' $\rho_k$ and to improve upon the factor of radial increase in~\ref{eq:methods:repeated-rnn-factor} so that this occurs less frequently and with less severity.
